# Exploratory Data Analysis - Red & White Wines 

#### by Darryl Ma
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(gridExtra)
library(dplyr)
library(GGally)
library(scales)
library(memisc)
library(pander)
```

```{r echo=FALSE, Load_the_Data}
# Load the data
red <- read.csv('wineQualityReds.csv')
white <- read.csv('wineQualityWhites.csv')

red['color'] <- 'red'
white['color'] <- 'white'

wines <- rbind(red, white)

# Re-numbering wines
wines$X <- c(0:6496)

# Creating new variables to convert quality into an ordered variable, 
# categorize wines into low, medium and high quality wines
wines$quality.ordered <- factor(wines$quality, ordered = T)
wines$quality.bucket <- with(wines, cut(wines$quality, 
                                        breaks=c(0, 5, 7, 10), ordered=TRUE))
wines$quality.category <- ifelse(wines$quality.bucket == '(0,5]', "low", 
                          ifelse(wines$quality.bucket == '(5,7]', 
                                 "medium", "high"))
```

# Background

The purpose of this project is to emulate the process of exploring a data set
and eventually extract meaningful insights. For this particular exercise, a 
database of wines was explored. For each entry of wine, various properties
such as acidity, density, alcoholic content and pH along with a subjective
quality rating of 0 to 10 were provided. I set two main questions to be
answered through this exploration:

1. What properties can be used to differentiate red and white wines?
2. Can the quality rating be estimated using a combination of properties?

Below is the source of the dataset being explored: 

P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. Modeling wine 
preferences by data mining from physicochemical properties. In Decision 
Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

### High Level Characteristics of Dataset

* Number of observations: 6,497
* Number of attributes: 13
    1. fixed.acidity -> most acids involved with wine where it is fixed or 
    nonvolatile (does not evaporate readily) [g/dm^3]
    2. volatile.acidity -> the amount of acetic acid in wine, which at  
    high levels can lead to an unpleasant, vinegar taste [g/dm^3]
    3. citric.acid -> found in small quantities, citric acid can add 
    'freshness' and flavor to wines [g/dm^3]
    4. reisdual.sugar -> the amount of sugar remaining after fermentation stops, 
    it's rare to find wines with less than 1 gram/liter and wines with greater 
    than 45 grams/liter are considered sweet [g/dm^3]
    5. chlorides -> the amount of salt in the wine [g/dm^3]
    6. free.sulfur.dioxide -> the free form of SO2 exists in equilibrium 
    between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents 
    microbial growth and oxidation of wine [mg/dm^3]
    7. total.sulfur.dioxide -> amount of free and bound forms of S02; in low 
    concentrations, SO2 is mostly undetectable in wine, but at free SO2 
    concentrations over 50 ppm, SO2 becomes evident in the nose and taste of 
    wine [mg/dm^3]
    8. density -> the density of wine is close to that of water depending 
    on the percent of alcohol and sugar content [g/cm^3]
    9. pH -> describes how acidic or basic a wine is on a scale from 0 
    (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
    10. sulphates -> a wine additive, which can contribute to sulfur dioxide gas 
    (S02) levels and acts as an antimicrobial and antioxidant [g/dm^3]
    11. alcohol -> the percent alcohol content of the wine [% by volume]
    12. quality -> output variable (based on sensory data) [score between 0 
    and 10]
    13. color -> indicates color of the wine [red/white]
    14. quality.bucket -> segregate wines into three quality buckets: (0,5],
    (5,7], and (7,10]
    15. quality.category -> segragate wines into three categories: low (0.5], 
    medium (5,7], and high (7,10]

* Structure of dataset:

```{r echo=FALSE}
str(wines)
```

* Summary of dataset:

```{r echo=FALSE}
pandoc.table(summary(wines))
```
  
# Univariate Plots Section

```{r echo=FALSE}
# Helper function to plot a single histogram
plotHistogram <- function(x, xlabel, bw, qstart, qend) {
  ggplot(data = wines, aes(x = x)) +
    geom_histogram(binwidth = bw) +
    xlab(xlabel) +
    xlim(quantile(x, qstart), quantile(x, qend))
}

# Helper function to plot a single bar plot
plotBar <- function(x, xlabel) {
  ggplot(data = wines, aes(x = x)) +
    geom_bar() +
    xlab(xlabel)
}

# Helper function to plot a histogram for white wines and a histogram for red
# wines on the same plot
plotRedWhiteDensity <- function(x, xlabel, bw, qstart, qend) {
  ggplot(data = wines, aes(x = x)) +
    geom_density(binwidth = bw, aes(fill = color, alpha = 0.5, adjust=0.5,
                                      y = ..density../sum(..density..)*100)) +
    xlab(xlabel) +
    xlim(quantile(x, qstart), quantile(x, qend)) +
    ylab("Percentage (%)") +
    scale_alpha_continuous(name = "alpha")
}
```

Below are univariate plots for each wine property: 

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotHistogram(wines$fixed.acidity, "Fixed Acidity", 0.1, 0, 1)
```

There is a high concentration of wines with fixed acidity around the median of 7 
g/dm^3 but there are some outliers that increases the mean to 7.2g/dm^3.

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotHistogram(wines$volatile.acidity, "Volatile Acidity", 0.01, 0, 1)
```

There appears to be some bimodalism with volatile acid, with a strong peak at 
0.25g/dm^3 and another weaker peak at 0.58g/dm^3. Also there are a few outliers 
that deviate quite significantly from the median and mean that reach a maximum
of 1.58 g/dm^3. At such high levels, one would assume that the wine would have
a strong vinegar taste. Is this an indication that the wine has gone bad?

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotHistogram(wines$citric.acid, "Citric Acid", 0.01, 0, 1)
```

In general, citric acid demonstrates a normal distribution, whereby median and 
mean are almost equal at 0.31g/dm^3 and 0.32g/dm^3, respectively. There is 
strange spike at 0.5g/dm^3.

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotHistogram(wines$residual.sugar, "Residual Sugar", 0.1, 0, 0.999)
```

The plot for residual sugar is right skewed where there are a significant 
number of outliers with higher sugar content. Most wines, however, are 
concentrated around 3g/dm^3 (median).The plot above shows that the variability
in sugar content in wines is quite wide.

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotHistogram(wines$chlorides, "Chlorides", 0.002, 0, 0.995)
```

There appears to be some bimodalism with chlorides, with a strong peak at 
0.04g/dm^3 and another weaker peak at 0.075g/dm^3.

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotHistogram(wines$free.sulfur.dioxide, "Free Sulfur Dioxide", 1, 0, 0.999)
```

The plot for free sulfur dioxide is right skewed where there are a significant 
number of outliers with higher concentrations of free sulfur dioxide. 
Most wines, however, are concentrated around 29mg/dm^3 (median). 

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotHistogram(wines$total.sulfur.dioxide, "Total Sulfur Dioxide", 3, 0, 0.999)
```

There appears to be some bimodalism with total sulfur dioxide, with a strong 
peak at 120mg/dm^3 and another weaker peak at 25mg/dm^3.

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotHistogram(wines$density, "Density", 0.0005, 0, 0.999)
```

There appears to be some bimodalism with density, with two almost equal peaks  
at 0.993g/cm^3 and another weaker peak at 0.996g/cm^3. There is one white wine 
with a very high density of 1.039g/cm^3, but this is the very same wine
with a residual sugar level of 65.8g/dm^3 which is roughly 12 times the 
average residual sugar level of 5.4g/dm^3. This may be a very specialized 
type of wine manufactured for a very niche market.

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotHistogram(wines$pH, "pH", 0.01, 0, 0.999)
```

pH exhibits a very normal distribution, whereby median and mean are almost 
equal at 3.21 and 3.22, respectively. 

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotHistogram(wines$sulphates, "Sulphates", 0.01, 0, 0.999)
```

The plot for sulpahtes is slightly right skewed where the mean at 0.53g/dm^3 
is slightly greater than the median at 0.51g/dm^3. Most wines, however, are 
concentrated around 0.51g/dm^3 (median). 

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotHistogram(wines$alcohol, "Alcohol", 0.1, 0, 0.999)
```

The plot for alcohol is right skewed with a peak at 9%. Given that the median 
is 10.3%, we can see that this ditribution does not follow a normal 
distribution.

```{r echo=FALSE, warning=FALSE, fig.width=7}
plotBar(wines$quality.ordered, "Quality")
```

Quality appears to be normally distributed with median equal to 6 and mean 
equal to 5.8.

I was slightly concerned that there were so many variables that exhibited 
bimodalism. Below are those list of variables: 

1. volatile acidity
2. citric acid
3. residual sugar
4. chlorides
5. free sulfur dioxide
6. total sulfur dioxide
7. density
8. alcohol 

However, I wondered whether this was simply because white and red wines differ 
for these particular properties. Suspecting this was the case, I re-plotted all
the graphs above, differentiating between red and white wines:

```{r echo=FALSE, warning=FALSE, fig.width=8, fig.height=12}
p1 <- plotRedWhiteDensity(wines$fixed.acidity, "Fixed Acidity", 0.1, 0, 0.99)
p2 <- plotRedWhiteDensity(wines$volatile.acidity, "Volatile Acidity", 
                          0.01, 0, 0.99)
p3 <- plotRedWhiteDensity(wines$citric.acid, "Citric Acid", 0.01, 0, 0.99)
p4 <- plotRedWhiteDensity(wines$residual.sugar, "Residual Sugar", 0.1, 0, 0.99)
p5 <- plotRedWhiteDensity(wines$chlorides, "Chlorides", 0.002, 0, 0.99)
p6 <- plotRedWhiteDensity(wines$free.sulfur.dioxide, "Free Sulfur Dioxide", 
                          1, 0, 0.99)
p7 <- plotRedWhiteDensity(wines$total.sulfur.dioxide, "Total Sulfur Dioxide", 
                          1, 0, 0.99)
p8 <- plotRedWhiteDensity(wines$density, "Density", 0.0001, 0, 0.99)
p9 <- plotRedWhiteDensity(wines$pH, "pH", 0.01, 0, 0.99)
p10 <- plotRedWhiteDensity(wines$sulphates, "Sulphates", 0.05, 0, 0.99)
p11 <- plotRedWhiteDensity(wines$alcohol, "Alcohol", 0.1, 0, 0.99)
p12 <- plotRedWhiteDensity(wines$quality, "Quality", 0.5, 0, 1)

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12, ncol = 2)
```

True enough, the plots above show that for almost all properties except 
fixed acidity, alcohol, and quality, there is a clear distinction between red 
and white wines, which accounts for the bimodalism in most cases. For example, 
red wines have a distinctively lower total sulphur dioxide concentration 
compared to white wines, which is the reason for the bimodalism when the 
datasets are observed together.

However even for the properties that are seemingly normally distributed, when 
the red and white wine samples are analyzed separate, properties such as pH are 
distinctively higher for red wines compared to white wines.

# Univariate Analysis

This dataset contains 6,497 observations (i.e. 6,497 different types of wines - 
1,599 red wines and 4,898 white wines). Each observation has 13 variables or 
properties that describe the characteristics of the wine. Of the 13 properties, 
11 are numeric, 1 is an integer, and 1 was a string variable. 3 additional 
variables were created: quality.ordered, quality.bucket and quality.category. 
This was to help analyze quality as an ordered variable as well as segregate
the wines into categories of low, medium and high quality.

Whilst investigating, I did find some bimodal distributions but it seems 
the bimodalism was due more in part to differences in properties between red and 
white wines rather than bimodalism inherent to the variables themselves. Further 
investigation showed that white and red wines are distinctively different for 
almost all properties except fixed acidity, alcohol, and quality. This is an 
important observation as it meant that any future analysis may require 
red and white wines to be analyzed separately.

Another observation was a noticeable spike in counts at the citric acid level 
of 0.5 for both red and white wines. A quick search on 
[Wikipedia](https://en.wikipedia.org/wiki/Acids_in_wine#Citric_acid) showed 
that  wine grapes naturally have an citric acidity of 0.05. Citric acid is often
artificially added to wines to boost the wine's acidity. Therefore the spike at 
0.5 may be particular to a certain brand of wine.

Ultimately, this univariate exploration helped me achieve my first goal, which 
was to identify properties that can be used to differentiate between red and 
white whites. Some key observations:

1. If you are looking for a sweet wine, you will have more chances of finding
what you are looking for if you purchase a white wine
2. White wines tend to be less dense (i.e. lighter) compared to red wines
3. White wines tend to be slightly more acidic than red wines
4. If you are unsure about the alcoholic content of a bottle of wine, 
it would not matter if it is red or white wine as the distribution of alcholic
content is similar for both types of wine.
 
To achieve the next goal, which was to identify a model that can 
estimate wine quality based on its properties, it is important to
get sufficient data points for low, medium and high quality wines. 
Unfortunately, based on the univariate exploration above, the lowest quality 
wine is rated at 3 and the highest quality wine is rated at 9, with 
50% of the wines being rated either 5 or 6, meaning there are fewer data 
points to help build the prediction model at higher and lower ratings.

# Bivariate Plots Section

```{r echo=FALSE}
# Helper function to plot a scatter plot given two variables
plotScatter <- function(x, xlabel, y, ylabel, qstart, qend) {
  ggplot(data = white, aes(x = x, y = y)) +
    geom_point(alpha = 0.1) +
    xlab(xlabel) +
    xlim(quantile(x, qstart), quantile(x, qend)) +
    ylab(ylabel) +
    ylim(quantile(y, qstart), quantile(y, qend)) +
    geom_smooth(method = 'lm', color = 'red') + 
    scale_color_brewer(type = 'div')
}

# Helper function to plot boxplots given two variables
plotBoxPlot <- function(x, xlabel, y, ylabel, qstart, qend) {
  ggplot(data=wines, aes(y = y, x = x)) + 
  geom_boxplot() +
    xlab(xlabel) +
    ylab(ylabel) +
    ylim(quantile(y, qstart), quantile(y, qend))
}
```

The best course of action when you do not know where to begin is to plot 
everything in the hopes that something catches your attention and helps narrow 
down the variables most relevant for your analysis. However, as higlighted 
before, because red and white wines are distinctively different for many of 
these properties, the pairings plot below only compares variables for 
white wines:

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=10}
ggpairs(data = subset(wines, wines$color == "white")) + 
        theme(text = element_text(size = 8))
```

Below are the variables that have a medium or high correlation coefficient 
assuming the following definitions:

* High correlation: 0.5 to 1.0 or -0.5 to -1.0
* Medium correlation: 0.3 to 0.5 or -0.3 to -0.5
* Low correlation: -0.3 to 0.3

Variable 1          | Variable 2           | Corr Coefficient | Correlation
----------------    | ---------------------| -----------------|------------
Residual Sugar      | Density              | 0.839            | High
Alcohol             | Density              | -0.78            | High
Total Sulfur Dioxide| Density              | 0.53             | High
Residual Sugar      | Alcohol              | -0.451           | Medium
Alcohol             | Total Sulfur Dioxide | -0.449           | Medium
pH                  | Fixed Acidity        | -0.426           | Medium
Residual Sugar      | Total Sulfur Dioxide | 0.401            | Medium

I have excluded correlation relationships with the variable X, relationships
between free and total sulfur dioxide as these two variables are essentially
measuring sulfur dioxide, as well as relationships with quality as these will
be addressed separately.

At first glance, it makes sense that density, residual sugar and alcohol have
a strong correlation. After all, density is a function of percent alcohol 
and sugar content. See plots below:

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=7, fig.height=10}
p1 <- plotScatter(white$residual.sugar, "Residual Sugar", white$density, 
                  "Density", 0, 0.99)
p2 <- plotScatter(white$alcohol, "Alcohol", white$density, "Density", 0, 0.99)

grid.arrange(p1,p2, ncol = 1)
```

Wine density increases as more sugar is added but decreases as alcholic 
content increases.

Another expected correlation is the negative relationship between fixed acid
and pH. That is, as fixed acidity content levels increase, the pH level
decreases or becomes more acidic. See plot below:

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=7}
p1 <- plotScatter(white$fixed.acidity, "Fixed Acidity", white$pH, "pH", 0, 1)

grid.arrange(p1, ncol = 1)
```

For quality, the following two variables had significant correlation:

* alcohol 
* density

**Note**: When analyzing quality, we considered both red and white wines 
together as there did not appear to be a clear disctinction between the wines 
in terms of quality.

```{r echo=FALSE, warning=FALSE, message=FALSE}
p1 <- plotBoxPlot(wines$quality.ordered, "Quality", wines$alcohol, 
                  "Alcohol", 0, 0.99)
p2 <- plotBoxPlot(wines$quality.ordered, "Quality", wines$density, 
                  "Density", 0, 0.99)

grid.arrange(p1,p2, ncol = 2)
```

In general, as alcoholic content increases, the quality rating increases. 
At a alcoholic levels of ~10%, however, the alcoholic content doesn't seem to 
influence the quality rating of the wine. As one might expect, because there 
is a negative relationship between percent alcohol content and density, as 
density decreases, quality of the wine increases.

But given the strong relationship between alcohol, density and quality,
I was surprised not to see a similar relationship between density, residual 
sugar and quality:

```{r echo=FALSE, warning=FALSE, message=FALSE}
p1 <- plotBoxPlot(wines$quality.ordered, "Quality", wines$residual.sugar, 
                  "Residual Sugar", 0, 0.99)

grid.arrange(p1, ncol = 2)
```

As shown above, a wine sweetness has little correlation to its quality.

# Bivariate Analysis

Some of the relationships observed were expected, and it was reassuring that
the data re-confirmed these relationship such as:

1. Postive correlation between density and residual sugar in white wines, 
which was the strongest relationship
2. Negative correlation between density and alcohol in white wines
3. Negative correlation between fixed acidity and pH in white wines

Other findings were new learnings for me as I did not expect to find 
a relationship between quality, alcohol and density:

1. Above an alcohol level of 10%, a wine is considered of higher quality as 
alcoholic content increases
2. Wines that are less dense are considered to be of higher quality
3. Sugar content does not factor significantly into how a wine is rated. 
At the same sugar level, quality ratings varied between 3 and 7.

One observation I found particularly interesting was that we found correlation
between fixed acidity and pH yet did not find a correlation between volatile
acidity and pH. I would have assumed that both fixed and volatile acidity 
would have had a strong negative correlation with pH.

# Multivariate Plots Section

The following plots try to explore if there is a distinction between various
wine qualities by looking at the relationship between alcoholic content 
and density:

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(data = wines, aes(x = alcohol, y = density)) +
    geom_point(aes(color = quality.ordered)) +
    xlim(quantile(wines$alcohol, 0), quantile(wines$alcohol, 0.99)) +
    ylim(quantile(wines$density, 0), quantile(wines$density, 0.99)) +
    geom_smooth(method = 'lm', aes(color = quality.ordered)) + 
    scale_color_brewer(type = 'seq', name = 'Quality') + 
  facet_wrap(~ color)
```

The separation of colors, where higher quality wines are concentrated on the 
right and lower quality wines are concentrated on the left, a behavior that is 
observed in both red and white wines, suggests that alcoholic content and 
density could be a good proxy to estimate wine quality. The following is a 
linear model of such an estimation:

```{r echo=FALSE, warning=FALSE, message=FALSE}
m1 <- lm(quality ~ alcohol, data = wines)
m2 <- update(m1, ~ . + density)
m3 <- update(m2, ~ . + fixed.acidity)
mtable(m1, m2, m3, sdigits = c(3))
```

To checked its precision, the model was used to predict quality at 95% 
confidence for a few test cases:

```{r echo=FALSE, Multivariate_Plots}
# Loading in test case
testData = read.csv('testData.csv')

# Create data frame of results 
modelEstimate = predict(m3, newdata = testData,
                        interval="prediction", level = .95)
modelEstimate <- data.frame(modelEstimate)

ggplot(data=modelEstimate, aes(x = testData$alcohol)) +
  geom_line(aes(y = modelEstimate$fit, color = 'Fit')) +
  geom_line(aes(y = modelEstimate$lwr, color = 'Lower Bound')) +
  geom_line(aes(y = modelEstimate$upr, color = 'Upper Bound')) +
  ylab("Predicted Quality") +
  xlab("Alcohol") +
  scale_color_discrete(name = 'Legend')
```

Below is the average difference between the lower/upper bounds of the 
95% confidence interval and best-fit prediction:

```{r echo=FALSE}
# Calcuate the mean different between the lower/upper bounds and fit
modelEstimate$diff_lwr <- modelEstimate$lwr - modelEstimate$fit
modelEstimate$diff_upr <- modelEstimate$upr - modelEstimate$fit
mean(modelEstimate$diff_lwr)
mean(modelEstimate$diff_upr)
```

From statistics above, it shows that the linear model can predict quality with
95% confidence of +/- 1.53. Meaning that if the model predicts a quality 
rating of 5.6, you can assume with 95% confidence that the actual quality value
is somewhere in between 4.1 and 7.1. Obviously this is not perfect, but if you 
are simply trying to gauge if the wine is of low, medium or high quality, then
it is probably safe to assume that it is of medium quality.

# Multivariate Analysis

The relationship between alcohol, density and quality was reconfirmed by 
plotting all three variables at once and observing the separation of qualities
by color. That is, wines with lower density and higher alcoholic content 
were rated higher and, inversely, wines with higher density and lower 
alcoholic content were rated lower. This charateristic was true of both 
red and white wines.

For this analysis a linear model was created using alcohol content, density, 
and fixed acidity to predict quality ratings. The r squared value of the 
predictive model was quite weak at 0.199, which rendered a 95% confidence 
interval of +/- 1.53 when predicting quality ratings. Nevertheless, if the 
objective of the model is to simply to differentiate between low, medium and 
high quality wines, this model should suffice.

------

# Final Plots and Summary

### Plot One

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=12}
p1 <- plotRedWhiteDensity(wines$volatile.acidity, "Volatile Acidity (g/dm^3)", 
                          0.01, 0, 0.99)
p2 <- plotRedWhiteDensity(wines$chlorides, "Chlorides (g/dm^3)", 0.002, 0, 0.99)
p3 <- plotRedWhiteDensity(wines$free.sulfur.dioxide, 
                          "Free Sulfur Dioxide (mg/dm^3)", 1, 0, 0.99)
p4 <- plotRedWhiteDensity(wines$total.sulfur.dioxide, 
                          "Total Sulfur Dioxide (mg/dm^3)", 1, 0, 0.99)
p5 <- plotRedWhiteDensity(wines$density, "Density (g/cm^3)", 0.0001, 0, 0.99)
p6 <- plotRedWhiteDensity(wines$pH, "pH", 0.01, 0, 0.99)
p7 <- plotRedWhiteDensity(wines$sulphates, "Sulphates (g/dm^3)", 0.05, 0, 0.99)

grid.arrange(p1,p2,p3,p4,p5,p6,p7, ncol = 2, top = "Red Vs White Wines")
```

### The Difference Between Red and White Wines 

The plots above demonstrate the distinct differences between red and white 
wines. In general, red whites tend to have higher volatile acidity, 
cloride content, density, pH levels, and sulphate content. Whilst white wines 
tend to have higher sulfur dioxide content.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = wines, aes(x = alcohol, y = density)) +
    geom_point(aes(color = quality.ordered)) +
    xlim(quantile(wines$alcohol, 0), quantile(wines$alcohol, 0.99)) +
    ylim(quantile(wines$density, 0), quantile(wines$density, 0.99)) +
    geom_smooth(method = 'lm', aes(color = quality.ordered)) +
    scale_alpha_continuous(name="alpha") +
    scale_color_brewer(type = 'seq', name = 'Quality') + 
    xlab("Alcohol (% of volume)") +
    ylab("Density (g/cm^3)") +
    labs(title = "Quality Vs Alcohol Vs Density") +
  facet_wrap(~ color)
```

### Quality Vs Alcohol Vs Density

The plot above demonstrates the relationship between quality, alcohol and 
density whereby quality rating are higher as alcohol content increases
and density decreases and, inversely, quality ratings are lower as alcohol 
content decreases and density increases. This behavior is apparent in both red 
and white wines.


### Plot Three
```{r echo=FALSE}
ggplot(data=modelEstimate, aes(x = testData$alcohol)) +
  geom_line(aes(y = modelEstimate$fit, color = 'Fit')) +
  geom_line(aes(y = modelEstimate$lwr, color = 'Lower Bound')) +
  geom_line(aes(y = modelEstimate$upr, color = 'Upper Bound')) +
  ylab("Quality") +
  xlab("Alcohol (% of volume)") +
  labs(title = "Predicting Quality") +
  scale_color_discrete(name = 'Legend')
```

### Accuracy of Linear Model

The linear model above shows predicted quality ratings along with the 
lower and upper bounds of the 95% confidence interval. The relatively large 
difference between the lower and upper bounds deomnstrates that the model is 
not particularly precise in predicting quality based on alcoholic content, 
density and fixed acidity values. However at the very least, it can be used to 
distinguish between low, medium and high quality wines.

------

# Reflection

At the start of this investigation, two goals were established:

1. To find properties that differentiate white and red wines
2. To predict quality rating based on a combination of wine properties 

For the first goal, we discovered that pH, sulphates, chlorides, volatile
acidity, free/total sulfur dioxide and density can all be used to differentiate
between red and white wines. 

For the second goal, a linear model was created to estimate quality based
on alcoholic content, density and fixed acidity. Though the r squared value of 
the model was weak at 0.199, the model showed that it could be somewhat useful 
in differentiating between low, medium and high quality wines. 

Given the inability to accurately estimate wine quality ratings, one may 
conclude that wine ratings are somewhat subjective in nature and are highly 
dependent on the personal tastes of the testers. Another concern to consider 
is the limitation of the variables provided. Perhaps other variables like 
year of production, wine brand, location where grapes were harvested, etc. 
have more of an influence over quality rating than the variables that we 
analyzed.

In conclusion, the practical takeaways from this exploration are as such:

1. If you want to appear like an amateur wine connoisseur, check the alcohol 
content label and test for density before giving an assessment
2. Make sure you provide a vague assessment (i.e. good, average, bad) 
3. If your assessment of the quality is way off, just attribute it to
preference in taste
